---
title: "The Predictiveness of Mental Health Diagnostic Groupings on High Cost Claim Frequency"
author: "John Ryan Kivela, MA"
subtitle: "*Northern Arizona Regional Behavioral Health Alliance <br> Flagstaff, AZ*"
date: '`r Sys.Date()`'
output: 
  html_document:
      theme: cerulean
      toc: yes
      toc_float: yes
      toc_depth: 2
      toc_collapsed: no
      code_folding: show
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r Load Packages and Libraries, echo=FALSE, message=FALSE, warning=FALSE}
# Load Packages and Libraries
library(tidyverse)
library(psych)
library(jtools)
library(kableExtra)
library(broom.mixed)
library(stargazer)
library(gtsummary)
library(gt)
```

```{r load data, echo=FALSE, message=FALSE}
ICD10_Groupings_Original <- read_csv("Data_Files/ICD10 BH Groupings.csv")
Claims_IPClaimsPerAdmission <- read_csv("Data_Files/20221231_FUH_ClaimsHx_FY20221 - R2.csv")  
InpatientCodeDescriptions <- read_csv("Data_Files/My2022_InpatientCodesDesc.csv")
ICD10DxAndGroupsCombined <- read_csv("Data_Files/ICD10andBHGroupingsCombined.csv")
```

```{r Create Objects, include=FALSE}
#Knitting Yarn
DateExtracted <- "12/31/22"

ClaimsAdjudicatedThrough <- "October 1, 2021 to September 30, 2022"

RegressionModel <- "Predicted High Cost Claim Score = `(F10-F19)` + `(F20-F29)` + `(F30-F39)` + `(F40-F48)` + `(F60-F69)` + `(F90-F98)` + `(F80-F89)` + `(F01-F09)`, Note: `(F50-F59)`# removed b/c contains only NA"

Provider_ShortName <- (c("CBI", "CPIH", "EHS", "LCBHC", "MMHC", 
                           "SHG", "SBH", "TGC", "Polara"))
# Wrangle the data
IPCodeDesciptions <- InpatientCodeDescriptions %>% 
  rename("svccode" = "InpatientCode")

Knitting <- data.frame(
                  DateExtracted,
                  ClaimsAdjudicatedThrough,
                  RegressionModel
                  )

```

```{r create_distribution_table, echo=FALSE, message=FALSE, warning=FALSE}
# Filter claims for only facility and Alliance Providers
Tab_DxAnalysis <- Claims_IPClaimsPerAdmission %>% 
  filter(svccode < 300) %>%  #filters to only facility claims
  filter(ra %in% Provider_ShortName) #filters to Alliance Providers

# Join claims table with My2022 IP Code Descriptions 
Tab_DxAnalysis <- left_join(x = Tab_DxAnalysis,
                            y = IPCodeDesciptions,
                            by = "svccode",
                            keep = TRUE)

# Join Tab_Dx table with ICD10 Diagnosis Grouping table
TestTab_DxAnalysis <- left_join(x = Tab_DxAnalysis,
                            y = ICD10DxAndGroupsCombined,
                            by = "primaryDiagnosis",
                            keep = TRUE)

# Select variables for analysis
TestTab_DxAnalysis <- TestTab_DxAnalysis %>% 
  select(ID,
         svcdate,
         primaryDiagnosis.x,
         svccode.x,
         units,
         calcnetpd,
         ra,
         DiagnosisGroupRange,
         Diagnosis_Group
         )
```

```{r descriptive_statistics, echo=FALSE, message=FALSE}
# Descriptive Statistics
Descriptive_Stats <- as.data.frame(round(describe(TestTab_DxAnalysis$calcnetpd), digits = 2))

# Create a Data frame for key descriptive statistics
Mean_CalcnetPd <- Descriptive_Stats$mean
SD_CalcnetPd <- Descriptive_Stats$sd
Outlier_Cutoff = Mean_CalcnetPd + (3*SD_CalcnetPd)
n_Stats <- Descriptive_Stats$n

# Remove Outliers
TestTabDx_Outliers <- TestTab_DxAnalysis %>% 
  filter(calcnetpd >= Outlier_Cutoff)

TestTabDx_NoOutliers <- TestTab_DxAnalysis %>% 
  filter(calcnetpd <= Outlier_Cutoff)

# Count the outliers
Outliers <- nrow(TestTabDx_Outliers)

# Run updated descriptive stats with no outliers. 
Descriptive_Stats2 <- as.data.frame(round(describe(TestTabDx_NoOutliers$calcnetpd), digits = 2))

# Create a Data frame for key descriptive statistics
Mean_CalcnetPd2 <- Descriptive_Stats2$mean
SD_CalcnetPd2 <- Descriptive_Stats2$sd
n_Stats2 <- Descriptive_Stats2$n

AllStats <- data.frame(Mean_CalcnetPd,
                       Mean_CalcnetPd2,
                       SD_CalcnetPd,
                       SD_CalcnetPd2,
                       Outliers,
                       Outlier_Cutoff,
                       n_Stats,
                       n_Stats2,
                       RegressionModel)

Allstats <- mutate(AllStats, as.numeric(n_Stats))

options(scipen = 999, show.signif.stars = TRUE)

AllStats <- AllStats %>% 
  mutate(Mean_CalcnetPd = scales::dollar(Mean_CalcnetPd)) %>% 
  mutate(Mean_CalcnetPd2 = scales::dollar(Mean_CalcnetPd2)) %>% 
  mutate(SD_CalcnetPd = scales::dollar(SD_CalcnetPd)) %>% 
  mutate(SD_CalcnetPd2 = scales::dollar(SD_CalcnetPd2)) %>% 
  mutate(Outlier_Cutoff = scales::dollar(Outlier_Cutoff)) 

```

```{r Regression_InitialRegression, echo=FALSE, message=FALSE}
# Regression Analysis Diagnosis Group Range by Calcnetpd
RegressionTest <- TestTabDx_NoOutliers %>%
  group_by(svcdate, DiagnosisGroupRange) %>%
  summarise(sum_calcnetpd = sum(calcnetpd, na.rm = FALSE)) %>%
  pivot_wider(names_from = DiagnosisGroupRange,
              values_from = sum_calcnetpd)

# Regression Analysis Diagnosis Group by Calcnetpd
# Create a dataframe for thhe sum of calcnetpd for the entire svcdate
SumCalcnetPd <- TestTabDx_NoOutliers %>% 
  group_by(svcdate) %>% 
  summarise(calcnetpd = sum(calcnetpd, na.rm = FALSE) )

# Join Sumcalcnetpd and regression test table
RegressionTest <- inner_join(x = RegressionTest,
                             y = SumCalcnetPd,
                             by = "svcdate")

# change any NA values to 0
RegressionTest[is.na(RegressionTest)] = 0

# run the regression
# plot(RegressionTest)

simple.regression <- lm(formula = calcnetpd ~ `(F10-F19)`
                        + `(F20-F29)` + `(F30-F39)`
                        + `(F40-F48)` + `(F60-F69)`
                        + `(F90-F98)` 
                        + `(F80-F89)`
                        # + `(F50-F59)`# removed b/c contains only NA
                        + `(F01-F09)`,
                        data = RegressionTest)

# summary(simple.regression)

```

```{r Regression_SecondaryRegression, echo=FALSE, message=FALSE}
# Secondary Regression Test
RegressionTest2 <- RegressionTest %>% 
  select(svcdate,
         `(F20-F29)`,
         `(F30-F39)`,
         calcnetpd)

simple.regression2 <- lm(formula = calcnetpd ~ `(F20-F29)` + `(F30-F39)`,
                        data = RegressionTest2)

# summary(simple.regression2)

```

# Abstract

This study is an analysis of adjudicated claims for Alliance ACO
Providers in the Northern Region of Arizona. The purpose of this
investigation is to identify areas of focus for program development and
clinical advancement by identifying predictive variables from the most
recent contract year, 2021-2022.

A predictive analysis of high cost claims and ICD10 mental health
groupings was conducted showing that *Schizophrenia, schizotypal,
delusional, and other non-mood psychotic disorders* (F20-F29) and *Mood
[affective] disorders* (F30-F39) are strong predictors of high cost
claims.

# Hypothesis

Recent feedback from Alliance ACO Providers, and the recent results of
the Alliance ACO Progress Report, support a closer analysis of the
factors impacting inpatient admissions for mental health conditions.
While the Alliance as a whole demonstrated an above average performance
for the 6 month baseline period, some individual providers failed to
meet the NCQA Mean for Follow-Up After Hospitalization for Mental
Illness (FUH)[^1].

[^1]: Kivela J.R., McMillian, J. & Tewa, V. (2023) Alliance ACO Progress
    Report, January 2023. *The Northern Arizona Regional Behavioral
    Health Authority*.

This study hypothesizes that different diagnostic groupings of mental
health disorders may have a greater impact on high cost claims. These
insights will help the Alliance develop data-driven interventions for
groups at greater risk of high cost claims.

# Method

## Data Management, transparency, and integrity

The Alliance collects referential data through direct query of curated
BCBSAZ hosted servers and databases. As such, the data has, at a
minimum, the level of security and validation offered by BCBSAZ
services.

Data is stored using Microsoft Office 365 cloud storage. Office 365 uses
several encryption protocols, including Transport Layer Security/Secure
Sockets Layer (TSL/SSL), Internet Protocol Security (IPSec), and
Advanced Encryption Standard (AES). Office 365 data encryption applies
to data at rest (data stored in One Drive for Business or SharePoint
libraries) and data in transit (mail messages, notes from online
meetings, or whenever a device is communicating within or between
servers[^2].

[^2]: Encryption in Microsoft 365 - Microsoft Purview (Compliance), 2022

### *Description of the data sample*

This sample includes adjudicated claims from BCBSAZ for the period of
`r Knitting$ClaimsAdjudicatedThrough`. The population includes
individuals with an adjudicated claim for a HEDIS My2022 qualifying
inpatient code associated with a HEDIS My2022 qualifying mental health
diagnosis[^3]. The sample is filtered to only include claims for BCBSAZ
AHCCCS Complete Care members with residential addresses in the Northern
Geographical Service Area. The sample is further differentiated by
including only contracted Alliance Providers.

[^3]: <https://www.ncqa.org/hedis/>

The sample size was `r AllStats$n_Stats`.

## Preparation of Data for Analysis

The data was extracted from the BCBSAZ data warehouse on
`r DateExtracted` using SQL Management Studio. The SQL Code is available
for review on request. After extraction, the data was reviewed for
errors, such as duplications or NULL values. The original data is stored
in encrypted folders. Prior to analysis the data was de-identified by
converting date of birth to age, converting date of service to a
month/year format, and creating an alternative ID in lieu of patient
name or BCBSAZ ID.

### Preparing the data in R

Data files were loaded to the model:

-   De-identified claims history from the Contract Year 2021 - 2022

-   Reference tables for:

    -   ICD10 codes

    -   HEDIS My 2022 qualifying inpatient codes[^4]

    -   ICD 10 code groupings for mental health which will be used in a
        regression analysis[^5].

[^4]: <https://www.ncqa.org/hedis/>

[^5]: ICD-10-CM official guidelines for coding and reporting. (n.d.).
    Retrieved January 18, 2023, from
    <https://www.cms.gov/files/document/fy-2022-icd-10-cm-coding-guidelines-updated-02012022.pdf>

```{r load_data, echo=FALSE, message=FALSE}
# load data
```

Reference objects were created and some data was renamed to assist in
data frame construction and analysis.

```{r createobjects, echo=FALSE, message=FALSE}
# create objects
```

A date frame was created from the claims history, which was joined with
ICD10 mental health grouping identifiers and inpatient service codes
based on primary diagnosis.

```{r distributiontable, echo=FALSE, message=FALSE}
# create distribution table
```

# Results

## Assessment of Cost Distribution for All Providers and MH Groups

A series of analyses were conducted to observe the overall distribution
of high cost claims according to Alliance Provider and mental health
diagnostic grouping.

### *Preliminary analysis*

First, a basic histogram for cost of claim was plotted to observe the
distribution of data. This provided a very clear visual indicator that
the distribution is negatively skewed to the right by a small number of
high cost outliers. Bin width for the histogram was set to \$1,000.

$~$

```{r preliminary_analyses_histogram, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(TestTab_DxAnalysis,
       aes(x = calcnetpd)) +
  geom_histogram(binwidth = 1000,
                 color="#000775", 
                 fill="#e1e7fa")+
  geom_vline(aes(xintercept = mean(calcnetpd)), 
             color ="#9a0138", 
             size=.5)+
  geom_text(x = 13919,
            y = 8000,
            label = "Mean = $13,919",
            color = "#9a0138",
            hjust  = -.05)+
    theme_grey()+
    theme(axis.text.y = element_text(size = 6),
          strip.text.y = element_text(angle = 360)
          )+
          # scale_y_continuous(labels = scales::percent)+
          labs(title = "Histogram: Preliminary Analysis of High Cost Claims",
               #subtitle = "subtitle",
               caption = "SOURCE: BCBSAZ Adjudicated Claims, Oct 2021 - Sep 2022")+
          xlab(label = "Cost per Claim (dollars)")+
          ylab(label = "Number of Claims")
          
```

$~$

The distribution of cost was further investigated by expanding the
histogram to create facets for each mental health grouping of ICD10
codes[^6]. These results further demonstrated the data being skewed to
the right by very few cases.

[^6]: ICD-10-CM official guidelines for coding and reporting. (n.d.).
    Retrieved January 18, 2023, from
    <https://www.cms.gov/files/document/fy-2022-icd-10-cm-coding-guidelines-updated-02012022.pdf>

In addition, it was noted that the distribution appears to be
concentrated in the psychotic disorders, and mood disorders groups,
(F20-F29) and (F30-F39).

$~$

```{r preliminary_analyses_histogram_mentalhealthgroupings, echo=FALSE, message=FALSE}
ggplot(TestTab_DxAnalysis,
       aes(x = calcnetpd)) +
  geom_histogram(
    binwidth = 1000,
    color="#000775", 
    fill="#e1e7fa"
    )+
    facet_grid(rows = vars(DiagnosisGroupRange))+
  geom_vline(aes(xintercept = mean(calcnetpd)), 
             color ="#9a0138", 
             size=.5)+
  geom_text(x = 13919,
            y = 8000,
            label = "Mean = $13,919",
            color = "#9a0138",
            hjust  = -.02)+
    theme_grey()+
    theme(axis.text.y = element_text(size = 6),
               strip.text.y = element_text(angle = 360)
                )+
             # scale_y_continuous(labels = scales::percent)+
    labs(title = "Histogram: Cost per Claim by Diagnostic Grouping",
         subtitle = "Using ICD-10 Mental Health DIagnosis Groupings",
         caption = "SOURCE: BCBSAZ Adjudicated Claims, Oct 2021 - Sep 2022")+
    xlab(label = "Cost per Claim (dollars)")+
    ylab(label = "Number of Claims")
```

$~$

Descriptive statistics were calculated to further observe the
distribution and assist in decision making about suppressing outlying
high cost claims.

$~$

```{r calc_descriptive_statistics1, echo=FALSE, message=FALSE}
# Descriptive Statistics

```

```{r Tab_DesscriptiveStat1, echo=FALSE, message=FALSE}
Descriptive_Stats %>%
  select(n, mean, sd, min, max, skew, kurtosis) %>% 
gt() %>% 
  tab_header(
      title = md("**Descriptive Statistics: Preliminary Analysis**"),
      subtitle = md("Adjudicated Claim Cost Distribution (October 2021 - September 2022)")
    ) %>% 
    opt_align_table_header(align = "left") %>% 
    fmt_number(columns = c(n),
               rows = everything(),
               decimals = 0,
               sep_mark = ",") %>% 
    fmt_number(columns = c(skew, kurtosis),
               rows = everything(),
               decimals = 3,
               sep_mark = ",") %>%
    fmt_currency (columns = c(mean, sd, min, max),
               rows = everything(),
               decimals = 0,
               sep_mark = ",") %>% 
    # fmt_percent(Percent,
    #             rows = everything(),
    #             decimals = 2,
    #             incl_space = TRUE) %>% 
    cols_width(n ~ px(100),
              mean ~ px(100),
              sd ~ px(100),
              sd ~ px(100)) %>% 
    tab_source_note(source_note = md("*SOURCE: BCBSAZ Adjudicated Claims, extracted 12/31/22*")) %>% 
    tab_style(
      style = list(
        cell_fill(color = "#e1e7fa"),
        cell_text(weight = "bold")
      ),
      locations = cells_column_labels(
        columns = everything()
      )
    ) %>% 
    opt_horizontal_padding(scale = 3)
```

$~$

The results of these statistics were: Mean =
`r AllStats$Mean_CalcnetPd`, and Standard Deviation =
`r AllStats$SD_CalcnetPd`. The statistics confirmed our visual
observation that there is a negative skew in the data (Skewness =
`r Descriptive_Stats$skew`); however the shape of the distribution is
also impacted by a high kurtosis (Kurtosis =
`r Descriptive_Stats$kurtosis`), or a *heavy upper tail*, due to extreme
outliers. 

It was therefore determined that outliers beyond 3 standard
deviations of the mean in either direction will be removed. The outlier
cutoff, defined as the Mean added to three times the Standard Deviation
(Cutoff Limit = M + (3 \* SD)), gives us a cutoff limit of +/-
`r AllStats$Outlier_Cutoff`.

There were a total of `r AllStats$Outliers` outliers removed. The final
sample size is `r AllStats$n_Stats2` claims.

### *Secondary Analysis*

The histograms were re-constructed utilizing the new data set with
outliers suppressed. Bin width was again set to \$1000.The new chart for
the entire sample reveals a more symmetrical distribution with a slight
skew to the left (Skewness = `r Descriptive_Stats2$skew`). 

$~$

```{r Plot_Histogram_NoOutliers, echo=FALSE, message=FALSE}
# Cost per claim histogram by Grouping for whole sample
ggplot(TestTabDx_NoOutliers,
       aes(x = calcnetpd)) +
  geom_histogram(
    binwidth = 1000,
    color="#000775", 
    fill="#e1e7fa")+
  geom_vline(aes(xintercept = mean(calcnetpd)),col="#9a0138",size=.5)+
  geom_text(x = 13912,
            y = 8000,
            label = "Mean = $13,912",
            color = "#9a0138",
            hjust  = -.05)+
  theme_grey()+
  theme(axis.text.y = element_text(size = 4),
             #strip.text.y = element_text(angle = 360)
              )+
  labs(title = "Histogram: Secondary Analysis, Cost per Claim (Outliers Supressed)",
      # subtitle = "subtitle",
       caption = "SOURCE: BCBSAZ Adjudicated Claims, Oct 2021 - Sep 2022")+
          xlab(label = "Cost per Claim (dollars)")+
          ylab(label = "Number of Claims")
```

$~$


Updated descriptive statistics were calculated, revealing a new Mean =
`r AllStats$Mean_CalcnetPd2`, and Standard Deviation =
`r AllStats$SD_CalcnetPd2`. The mean was not greatly impacted by the
removal of the outlying values. The kurtosis of the distribution was
relieved by the removal of the outliers (Kurtosis =
`r Descriptive_Stats2$kurtosis`).

Altogether, the characteristics of the distribution indicate that this
data is acceptable to use for predictive analyses.

$~$

```{r Tab_DescriptiveStats2_NoOutliers, echo=FALSE, message=FALSE}
Descriptive_Stats2 %>%  
  select(n, mean, sd, min, max, skew, kurtosis) %>% 
gt()%>% 
  tab_header(
      title = md("**Descriptive Statistics: Secondary Analysis**"),
      subtitle = md("Outliers Removed: Adjudicated Claim Cost Distribution")
    ) %>% 
    opt_align_table_header(align = "left") %>% 
    fmt_number(columns = c(n),
               rows = everything(),
               decimals = 0,
               sep_mark = ",") %>% 
    fmt_number(columns = c(skew, kurtosis),
               rows = everything(),
               decimals = 3,
               sep_mark = ",") %>%
    fmt_currency (columns = c(mean, sd, min, max),
               rows = everything(),
               decimals = 0,
               sep_mark = ",") %>% 
    # fmt_percent(Percent,
    #             rows = everything(),
    #             decimals = 2,
    #             incl_space = TRUE) %>% 
    cols_width(n ~ px(100),
              mean ~ px(100),
              sd ~ px(100),
              sd ~ px(100)) %>% 
    tab_source_note(source_note = md("*SOURCE: BCBSAZ Adjudicated Claims, extracted 12/31/22*")) %>% 
    tab_style(
      style = list(
        cell_fill(color = "#e1e7fa"),
        cell_text(weight = "bold")
      ),
      locations = cells_column_labels(
        columns = everything()
      )
    ) %>% 
    opt_horizontal_padding(scale = 3)
```

$~$

An expanded histogram was constructed investigating facets of the total
cost of services including mental health ICD10 code groupings, and
responsible Alliance Provider. These results again appear to show a
concentration around (F20-F29) and (F30-F39). While the plot below is
somewhat complex, the expanded histogram provides additional information
concerning the frequency of high cost claims for Alliance Providers.

$~$

```{r Plot_ExpandedHistogram_NoOutliers, echo=FALSE, message=FALSE}
### Cost histogram by ra
ggplot(TestTabDx_NoOutliers,
       aes(x = calcnetpd)) +
  geom_histogram(
    binwidth = 1000,
    color="#000775", 
    fill="#e1e7fa"
  )+
  facet_grid(cols = vars(ra),
             rows = vars(DiagnosisGroupRange))+
  geom_vline(aes(xintercept = mean(calcnetpd)),col="#9a0138", alpha = .1, size=.5)+
  geom_text(x = 13912,
            y = 8000,
            label = "Mean = $13,912",
            color = "#9a0138",
            hjust  = -.05)+
  theme_grey()+
  theme(axis.text.y = element_text(size = 6),
        axis.text.x = element_text(angle = 60,
                                   hjust = 1,
                                   size = 8),
        strip.text.y = element_text(angle = 360)
        )+
  labs(title = "Histogram: Cost per Claim by Diagnostic Grouping (Outliers Suppressed)",
       subtitle = "Using ICD-10 Mental Health DIagnosis Groupings",
       caption = "Data Source")+
  xlab(label = "Cost per Claim (dollars)")+
          ylab(label = "Number of Claims")

```

## Regression Analysis

A multiple linear regression was conducted to test if ICD10 diagnostic
groupings for mental health significantly predicted the occurrence of
high cost claims. Preliminary analysis of the data reveals a
distribution that appears to be more heavily concentrated in the
psychotic disorders and mood disorders, (F20-F29) and (F30-F39)
groupings. 

The purpose of this test is to confirm if these groups are
significant predictors of high cost claims.

### *Preliminary regression analysis*

```{r Regression_InitialRegression1, echo=FALSE, message=FALSE}
# Regression Analysis Diagnosis Group Range by Calcnetpd

```

The results of the multiple linear regression reveal a significant
effect of diagnosis group on the occurrence of high cost claims overall
(*F*(8,3)=16.39, *p*=0.02), with (*R\^2* = 0.918).

This suggests that about **92% of the variation** is predicted by
diagnostic grouping. (F20-F29) and (F30-F39), psychotic disorders, and
mood disorders, were the highest predictors, with significance levels of
*p*=0.043 and *p*=0.013, respectively.

$~$

```{r Tab_Regression_InitialRegression, echo=FALSE, message=FALSE}
tbl_regression(simple.regression, intercept = TRUE) %>% 
  as_gt() %>% 
    tab_header(
          title = md("**Linear Regression: High Cost Claims Predicted by MH Grouping**"),
          subtitle = md("*Initial regression includes all groupings*")
        ) %>%
    opt_align_table_header(align = "left") %>%
    tab_source_note(source_note = md("*SOURCE: BCBSAZ Adjudicated Claims, extracted 12/31/22*")) %>% 
        tab_style(
          style = list(
            cell_fill(color = "#e1e7fa"),
            cell_text(weight = "bold")
          ),
          locations = cells_column_labels(
            columns = everything()
          )
        )

```

$~$

The forest plot below provides a visual way to observe the significance
of predictive variables. For each grouping the horizontal line
represents the confidence interval. This demonstrates a range of values
that we can be 95% certain contain the predicted true value. The
vertical line represents the value at which there is no significant
effect.

If the confidence interval line crosses the line of null effect, then
the result is not significant. (F20-29) and (F30-F39) are seen here as
significant predictors of high cost claims.

$~$

```{r Plot_Forest_Regression_InitialRegression, echo=FALSE, message=FALSE}
# Forest Plots 
plot_summs(simple.regression, plot.distributions = TRUE, inner_ci_level = .9)+
  theme_grey()+
  geom_vline(aes(xintercept = 0), 
             color ="#9a0138", 
             size=.5)+
  labs(title = "Effect level of Mental HealtH Groupings on High Cost Claims",
       subtitle = "*Using ICD-10 Mental Health DIagnosis Groupings*",
       caption = "SOURCE: BCBSAZ Adjudicated Claims, extracted 12/31/22")+
  xlab(label = "Level of Effect (*Beta*)")+
  annotate("text", x = 1, y = 8.5, label = "Line of Null Effect")

```

$~$

### Predictive grouping regression analysis

```{r Regression_SecondaryRegression1, echo=FALSE, message=FALSE}
# Secondary Regression Test

```

The multiple regression equation was updated to focus on the significant
predictive groupings (F20-F29) and (F30-F39). Overall, the updated
regression revealed a stronger statistically significant relationship
(*F*(2,9)=49.55, *p*\<0.001), with (*R\^2* = 0.898) when compared to the
original regression.

This highly significant result suggests that about **90% of the
variation** in high cost claim occurrence is predicted by ***these two
groupings alone***!. 

The value "$\beta$" represent the slope of the
regression line through the data. The level of significance was set
to "*p*\<.05", or a 95% Confidence Interval.

It was observed that (F20-F29) was a significant predictor of high cost
claims ($\beta$=1.2, *p*\<.001).

It was observed that (F30-F39) was a significant predictor of high cost
claims ($\beta$=1.0, *p*\<.001).

$~$

```{r Tab_Regression_SecondaryRegression, echo=FALSE, message=FALSE}
tbl_regression(simple.regression2, intercept = TRUE) %>% 
  as_gt() %>% 
    tab_header(
          title = md("**Linear Regression: High Cost Claims Predicted by MH Grouping**"),
          subtitle = md("*Non-predictive groupings suppressed*")
        ) %>%
    opt_align_table_header(align = "left") %>%
        tab_source_note(source_note = md("*SOURCE: BCBSAZ Adjudicated Claims, extracted 12/31/22*")) %>% 
        tab_style(
          style = list(
            cell_fill(color = "#e1e7fa"),
            cell_text(weight = "bold")
          ),
          locations = cells_column_labels(
            columns = everything()
          )
        )

```

$~$

By suppressing the non-predictive groupings, the significance of the
predictive value of (F20-F29) and (F30-F39) is more visible. The forest
plot below demonstrates the distribution of values contained within a
95% confidence interval for the 2 respective groupings. The
distributions are closely contained within the confidence interval, indicating that each distribution is statistically significant.

$~$

```{r Plot_Forest_Regression_SecondaryRegression1, echo=FALSE, message=FALSE}
# secondary regression
plot_summs(simple.regression2, plot.distributions = TRUE, inner_ci_level = .9)+
  theme_grey()+
  geom_vline(aes(xintercept = 0), 
             color ="#9a0138", 
             size=.5)+
  labs(title = "Effect level of Predictive MH Groupings on High Cost Claims)",
       subtitle = "Using ICD-10 Mental Health DIagnosis Groupings",
       caption = "SOURCE: BCBSAZ Adjudicated Claims, extracted 12/31/22")+
  xlab(label = "Level of Effect (Beta)")+
  annotate("text", x = 0, y = 2.6, 
           label = "Line of Null Effect",
           hjust = -0.1)
  
```

$~$

# Discussion

## Regression Analysis

The results of the analysis ***confirmed that the mental health
diagnostic group of the patient is a strong predictor*** for the
occurrence of high cost claims for inpatient residential services. In
particular, Schizophrenia, schizotypal, delusional, and other non-mood
psychotic disorders, and Mood [affective] disorders, (F20-F29) and
(F30-F39) respectively, were the highest predictors of all groupings,
accounting for ***90% of the variance*** across over 22,000 adjudicated
claims for a period of 1 year.

# Conclusion

## Why is this important?

Significant predictors of behaviors, services, and claim costs can be
hidden among the immense volume of data and clinical work that is
required to provide quality behavioral health services. It is common
knowledge in the field that some disorders tend to occur more frequently
than others in inpatient care. This research allows us to put a name,
and a significance level to the **top predictors**.

These results also helps us understand the power of psychotic disorders,
and mood disorders to predict higher cost. These 2 diagnostic groups,
out of 9, are responsible for **at least 90%** of the variance in our
sample.

These results provide evidence based decision making power to address
the clinical and administrative needs of members living with psychotic
and mood disorders. It also provides the ability to estimate performance
targets for reduction in inpatient service costs.

## Suggestions for Action

Further investigation into the inpatient experience of members with
predictive diagnosis is indicated. Areas of emphasis may include:

-   The impact of diagnostic groupings on length of stay

-   The impact of diagnostic groupings on re-admission

-   Frequency and duration of inpatient stays by facility

Additional training could be provided to enable the work force to
provide care to individuals with disorders that are highly predictive of
higher cost claims. Collaboration with health medical management
resources could also help to ensure clear and effective communication
between the health plan and Alliance Providers.
